Studies of quantum dots

### 2017-01-11 Manuscript/paper.tex
Update spEnergiesQD/Manuscript/paper.tex

\begin{figure}%[hbtp]
     \begin{center}
        \subfigure[Results for $N=6$ and $\omega=1.0$ ]{
            \label{fig:N6hw1}
            \includegraphics[width=0.38\textwidth]{figures/6parthw1.pdf}
        }
        \subfigure[Results for $N=6$ and $\omega=0.5$ ]{
           \label{fig:N6hw05}
           \includegraphics[width=0.38\textwidth]{figures/6parthw05.pdf}
        }\\ %  ------- End of the first row ----------------------%
        \subfigure[Results for $N=6$ and $\omega=0.28$ ]{
            \label{fig:N6hw028}
            \includegraphics[width=0.38\textwidth]{figures/6parthw028.pdf}
        }
        \subfigure[Results for $N=6$ and $\omega=0.1$ ]{
            \label{fig:N6hw01}
            \includegraphics[width=0.38\textwidth]{figures/6parthw01.pdf}
        }
    \end{center}
    \caption{Comparison of our IM-SRG(2) ground state energies (SRG)
      for circular quantum dots for $N=6$ and $\omega=1.0$
       The results are compared with diffusion Monte Carlo (DMC),
      coupled cluster at the level of singles and doubles (CCSD), full
      configuration interaction (FCI) and Hartree-Fock calculations
      (HF) as functions of the number of major oscillator shells
      $R$. A harmonic oscilaltor basis has been used, with an
      unrenormalized Coulomb repulsion and the White generator. All
      energies in atomic units.}
   \label{fig:N6}
\end{figure}

\begin{figure}%[hbtp]
     \begin{center}
        \subfigure[Results for $N=12$ and $\omega=1.0$ ]{
            \label{fig:N12hw1}
            \includegraphics[width=0.38\textwidth]{figures/12parthw1.pdf}
        }
        \subfigure[Results for $N=12$ and $\omega=0.5$ ]{
           \label{fig:N12hw05}
           \includegraphics[width=0.38\textwidth]{figures/12parthw05.pdf}
        }\\ %  ------- End of the first row ----------------------%
        \subfigure[Results for $N=12$ and $\omega=0.28$ ]{
            \label{fig:N12hw028}
            \includegraphics[width=0.38\textwidth]{figures/12parthw028.pdf}
        }
        \subfigure[Results for $N=12$ and $\omega=0.1$ ]{
            \label{fig:N12hw01}
            \includegraphics[width=0.38\textwidth]{figures/12parthw01.pdf}
        }
    \end{center}
    \caption{Comparison of our IM-SRG(2) ground state energies (SRG)
      for circular quantum dots for $N=12$ and different values of
      $\omega$. The results are compared with diffusion Monte
      Carlo (DMC), coupled cluster at the level of singles and doubles
      (CCSD), full configuration interaction (FCI) and Hartree-Fock
      calculations (HF) as functions of the number of major oscillator
      shells $R$. A harmonic oscilaltor basis has been used, with an
      unrenormalized Coulomb repulsion and the White generator. All
      energies in atomic units.}
   \label{fig:N12}
\end{figure}

\begin{figure}%[hbtp]
     \begin{center}
        \subfigure[Results for $N=20$ and $\omega=1.0$ ]{
            \label{fig:N20hw1}
            \includegraphics[width=0.38\textwidth]{figures/20parthw1.pdf}
        } \subfigure[Results for $N=20$ and $\omega=0.5$ ]{
           \label{fig:N20hw05}
           \includegraphics[width=0.38\textwidth]{figures/20parthw05.pdf}
        }\\ % ------- End of the first row ----------------------%
        \subfigure[Results for $N=20$ and $\omega=0.28$ ]{
            \label{fig:N20hw028}
            \includegraphics[width=0.38\textwidth]{figures/20parthw028.pdf}
        } \subfigure[Results for $N=20$ and $\omega=0.1$ ]{
            \label{fig:N20hw01}
            \includegraphics[width=0.38\textwidth]{figures/20parthw01.pdf}
        }
    \end{center}
    \caption{Comparison of our IM-SRG(2) ground state energies (SRG)
      for circular quantum dots for $N=20$ and different values of
      $\omega$. The results are compared with diffusion Monte
      Carlo (DMC), coupled cluster at the level of singles and doubles
      (CCSD), full configuration interaction (FCI) and Hartree-Fock
      calculations (HF) as functions of the number of major oscillator
      shells $R$. A harmonic oscilaltor basis has been used, with an
      unrenormalized Coulomb repulsion and the White generator. All
      energies in atomic units.}
   \label{fig:N20}
\end{figure}

\begin{figure}%[hbtp]
     \begin{center}
        \subfigure[Results for $N=30$ and $\omega=1.0$ ]{
            \label{fig:N30hw1}
            \includegraphics[width=0.38\textwidth]{figures/30parthw1.pdf}
        } \subfigure[Results for $N=30$ and $\omega=0.5$ ]{
           \label{fig:N30hw05}
           \includegraphics[width=0.38\textwidth]{figures/30parthw05.pdf}
        }\\ % ------- End of the first row ----------------------%
        \subfigure[Results for $N=30$ and $\omega=0.28$ ]{
            \label{fig:N30hw028}
            \includegraphics[width=0.38\textwidth]{figures/30parthw028.pdf}
        } \subfigure[Results for $N=30$ and $\omega=0.1$ ]{
            \label{fig:N30hw01}
            \includegraphics[width=0.38\textwidth]{figures/30parthw01.pdf}
        }
    \end{center}
    \caption{Comparison of our IM-SRG(2) ground state energies (SRG)
      for circular quantum dots for $N=30$ and different values of
      $\omega$. The results are compared with diffusion Monte
      Carlo (DMC), coupled cluster at the level of singles and doubles
      (CCSD), full configuration interaction (FCI) and Hartree-Fock
      calculations (HF) as functions of the number of major oscillator
      shells $R$. A harmonic oscilaltor basis has been used, with an
      unrenormalized Coulomb repulsion and the White generator. All
      energies in atomic units.}
   \label{fig:N30}
\end{figure}

\begin{figure}%[hbtp]
     \begin{center}
        \subfigure[Results for $N=42$ and $\omega=1.0$ ]{
            \label{fig:N42hw1}
            \includegraphics[width=0.38\textwidth]{figures/42parthw1.pdf}
        } \subfigure[Results for $N=42$ and $\omega=0.5$ ]{
           \label{fig:N42hw05}
           \includegraphics[width=0.38\textwidth]{figures/42parthw05.pdf}
        }\\ % ------- End of the first row ----------------------%
        \subfigure[Results for $N=42$ and $\omega=0.28$ ]{
            \label{fig:N42hw028}
            \includegraphics[width=0.38\textwidth]{figures/42parthw028.pdf}
        } \subfigure[Results for $N=42$ and $\omega=0.1$ ]{
            \label{fig:N42hw01}
            \includegraphics[width=0.38\textwidth]{figures/42parthw01.pdf}
        }
    \end{center}
    \caption{Comparison of our IM-SRG(2) ground state energies (SRG)
      for circular quantum dots for $N=42$ and different values of
      $\omega$. The results are compared with diffusion Monte
      Carlo (DMC), coupled cluster at the level of singles and doubles
      (CCSD), full configuration interaction (FCI) and Hartree-Fock
      calculations (HF) as functions of the number of major oscillator
      shells $R$. A harmonic oscilaltor basis has been used, with an
      unrenormalized Coulomb repulsion and the White generator. All
      energies in atomic units.}
   \label{fig:N42}
\end{figure}

\begin{table}
\begin{center}
  \caption{Ground state energy (in atomic units) of a circular quantum dot, with $N$ particles and oscillator frequency $\omega$. The IM-SRG(2) results are given for $R=20$ shells (except for the $N=56$ results which are for $R=18$), employing a Hartree-Fock basis, a bare Coulomb interaction and the White generator. For $N=42$ and $\omega=0.1$ results are presented up to $R=22$ major shells. The results labelled IM-SRG(2)$_{R\rightarrow \infty}$ are obtained using the extrapolation formula of [[equation?]].}
\label{tab:SRGDMC-R20}
\begin{tabular}{SSSSS}
  \toprule
  {$N$} & {$\omega$} & {IM-SRG(2)} & {IM-SRG(2)$_{R\rightarrow \infty}$} & {DMC} \\
  \midrule
  6 & 1.0 & 20.1583 & 20.1468   & 20.1593(1) \\
      & 0.5&11.7690  & 11.7631 &11.7848(1) \\
      & 0.28& 7.5713 &   7.5695 &7.6002(1) \\
      & 0.1& 3.4962  &  3.4977 &3.5539(1)\\
  \midrule
  12&1.0 & 65.7280 & 65.6990  & 65.7001(1) \\
      &0.5 &39.1631 &  39.1473 & 39.1596(1)\\
      &  0.28&25.6212 &  25.6140& 25.6358(1)\\
      &  0.1&12.2223 &  12.2234 & 12.2698(1)\\
  \midrule
  20& 1.0&155.9737 & 155.9130  & 155.8822(1)\\
      &  0.5&93.9219 &   93.8853 &93.8752(1)\\
      &  0.28&61.94362 & 61.9240 &61.9268(1) \\
      &  0.1&29.95263 &29.9565  & 29.9779(1)\\
  \midrule
  30& 1.0& 308.7673 & 308.6053 & 308.5627(2) \\
      & 0.5& 187.1671& 186.8340 &187.0426(2)\\
      & 0.28&124.0410 & 123.7171 & 123.9683(2) \\
      & 0.1& 60.4319& 60.1158 & 60.4205(2)\\
  \midrule
  42& 1.0&543.3398 & 543.3304 & 542.9428(8) \\
      & 0.5&330.8885 & 330.8425 & 330.6306(2)\\
      & 0.28&220.0226 & 219.9534y &219.8426(2) \\
      & 0.1& 107.7851&  106.8074& 107.6389(2)\\
  \midrule
  56& 1.0& 880.4159 &  & 879.3986(6) \\
      & 0.5& 538.8199 & & 537.353(2)\\
      & 0.28&360.83689  &  & 358.145(2) \\
  \bottomrule
\end{tabular}
\end{center}
\end{table}

@@

@@

There are a few cases where the IM-SRG over-corrects the result, leading to a ground state energy that is lower than the exact result.  This is not unexpected given that, unlike Hartree-Fock, IM-SRG is non-variational in the presence of operator truncations that cause small unitarity violations.  Such behavior tends to occur when the number of particles is few, or the frequency is very low (high correlation).  However, MPPT does not do much better in the high frequency regimes as it often converges slower than IM-SRG.

@@

\subsection{Slow convergence}

>>>>

\begin{figure}
  \centering
  \subfigure[Results for $N=6$ and $\omega=1.0$]{
    \label{fig:N6hw1}
    \includegraphics[width=0.38\textwidth]{figures/6parthw1.pdf}
  }
  \subfigure[Results for $N=6$ and $\omega=0.5$]{
    \label{fig:N6hw05}
    \includegraphics[width=0.38\textwidth]{figures/6parthw05.pdf}
  } \\
  \subfigure[Results for $N=6$ and $\omega=0.28$]{
    \label{fig:N6hw028}
    \includegraphics[width=0.38\textwidth]{figures/6parthw028.pdf}
  }
  \subfigure[Results for $N=6$ and $\omega=0.1$]{
    \label{fig:N6hw01}
    \includegraphics[width=0.38\textwidth]{figures/6parthw01.pdf}
  }
  \caption{Comparison of our IM-SRG(2) ground state energies (SRG) for circular quantum dots for $N=6$ and $\omega=1.0$ The results are compared with diffusion Monte Carlo (DMC), coupled cluster at the level of singles and doubles (CCSD), full configuration interaction (FCI) and Hartree-Fock calculations (HF) as functions of the number of major oscillator shells $R$. A harmonic oscilaltor basis has been used, with an unrenormalized Coulomb repulsion and the White generator. All energies in atomic units.}
  \label{fig:N6}
\end{figure}

\begin{figure}
  \centering
  \subfigure[Results for $N=12$ and $\omega=1.0$]{
    \label{fig:N12hw1}
    \includegraphics[width=0.38\textwidth]{figures/12parthw1.pdf}
  }
  \subfigure[Results for $N=12$ and $\omega=0.5$]{
    \label{fig:N12hw05}
    \includegraphics[width=0.38\textwidth]{figures/12parthw05.pdf}
  } \\
  \subfigure[Results for $N=12$ and $\omega=0.28$]{
    \label{fig:N12hw028}
    \includegraphics[width=0.38\textwidth]{figures/12parthw028.pdf}
  }
  \subfigure[Results for $N=12$ and $\omega=0.1$]{
    \label{fig:N12hw01}
    \includegraphics[width=0.38\textwidth]{figures/12parthw01.pdf}
  }
  \caption{Comparison of our IM-SRG(2) ground state energies (SRG) for circular quantum dots for $N=12$ and different values of $\omega$. The results are compared with diffusion Monte Carlo (DMC), coupled cluster at the level of singles and doubles (CCSD), full configuration interaction (FCI) and Hartree-Fock calculations (HF) as functions of the number of major oscillator shells $R$. A harmonic oscilaltor basis has been used, with an unrenormalized Coulomb repulsion and the White generator. All energies in atomic units.}
  \label{fig:N12}
\end{figure}

\begin{figure}
  \centering
  \subfigure[Results for $N=20$ and $\omega=1.0$]{
    \label{fig:N20hw1}
    \includegraphics[width=0.38\textwidth]{figures/20parthw1.pdf}
  } \subfigure[Results for $N=20$ and $\omega=0.5$]{
    \label{fig:N20hw05}
    \includegraphics[width=0.38\textwidth]{figures/20parthw05.pdf}
  } \\
  \subfigure[Results for $N=20$ and $\omega=0.28$]{
    \label{fig:N20hw028}
    \includegraphics[width=0.38\textwidth]{figures/20parthw028.pdf}
  } \subfigure[Results for $N=20$ and $\omega=0.1$]{
    \label{fig:N20hw01}
    \includegraphics[width=0.38\textwidth]{figures/20parthw01.pdf}
  }
  \caption{Comparison of our IM-SRG(2) ground state energies (SRG)
    for circular quantum dots for $N=20$ and different values of
    $\omega$. The results are compared with diffusion Monte
    Carlo (DMC), coupled cluster at the level of singles and doubles
    (CCSD), full configuration interaction (FCI) and Hartree-Fock
    calculations (HF) as functions of the number of major oscillator
    shells $R$. A harmonic oscilaltor basis has been used, with an
    unrenormalized Coulomb repulsion and the White generator. All
    energies in atomic units.}
  \label{fig:N20}
\end{figure}

\begin{figure}
  \centering
  \subfigure[Results for $N=30$ and $\omega=1.0$]{
    \label{fig:N30hw1}
    \includegraphics[width=0.38\textwidth]{figures/30parthw1.pdf}
  } \subfigure[Results for $N=30$ and $\omega=0.5$]{
    \label{fig:N30hw05}
    \includegraphics[width=0.38\textwidth]{figures/30parthw05.pdf}
  } \\
  \subfigure[Results for $N=30$ and $\omega=0.28$]{
    \label{fig:N30hw028}
    \includegraphics[width=0.38\textwidth]{figures/30parthw028.pdf}
  } \subfigure[Results for $N=30$ and $\omega=0.1$]{
    \label{fig:N30hw01}
    \includegraphics[width=0.38\textwidth]{figures/30parthw01.pdf}
  }
  \caption{Comparison of our IM-SRG(2) ground state energies (SRG)
    for circular quantum dots for $N=30$ and different values of
    $\omega$. The results are compared with diffusion Monte
    Carlo (DMC), coupled cluster at the level of singles and doubles
    (CCSD), full configuration interaction (FCI) and Hartree-Fock
    calculations (HF) as functions of the number of major oscillator
    shells $R$. A harmonic oscilaltor basis has been used, with an
    unrenormalized Coulomb repulsion and the White generator. All
    energies in atomic units.}
  \label{fig:N30}
\end{figure}

\begin{figure}
  \centering
  \subfigure[Results for $N=42$ and $\omega=1.0$]{
    \label{fig:N42hw1}
    \includegraphics[width=0.38\textwidth]{figures/42parthw1.pdf}
  } \subfigure[Results for $N=42$ and $\omega=0.5$]{
    \label{fig:N42hw05}
    \includegraphics[width=0.38\textwidth]{figures/42parthw05.pdf}
  } \\
  \subfigure[Results for $N=42$ and $\omega=0.28$]{
    \label{fig:N42hw028}
    \includegraphics[width=0.38\textwidth]{figures/42parthw028.pdf}
  } \subfigure[Results for $N=42$ and $\omega=0.1$]{
    \label{fig:N42hw01}
    \includegraphics[width=0.38\textwidth]{figures/42parthw01.pdf}
  }
  \caption{Comparison of our IM-SRG(2) ground state energies (SRG) for circular quantum dots for $N=42$ and different values of $\omega$. The results are compared with diffusion Monte Carlo (DMC), coupled cluster at the level of singles and doubles (CCSD), full configuration interaction (FCI) and Hartree-Fock calculations (HF) as functions of the number of major oscillator shells $R$. A harmonic oscilaltor basis has been used, with an unrenormalized Coulomb repulsion and the White generator. All energies in atomic units.}
  \label{fig:N42}
\end{figure}

\begin{table}
  \centering
  \caption{Ground state energy (in atomic units) of a circular quantum dot, with $N$ particles and oscillator frequency $\omega$. The IM-SRG(2) results are given for $R=20$ shells (except for the $N=56$ results which are for $R=18$), employing a Hartree-Fock basis, a bare Coulomb interaction and the White generator. For $N=42$ and $\omega=0.1$ results are presented up to $R=22$ major shells. The results labelled IM-SRG(2)$_{R\rightarrow \infty}$ are obtained using the extrapolation formula of [[equation?]].}
  \label{tab:SRGDMC-R20}
  \begin{tabular}{SSSSS}
    \toprule
    {$N$} & {$\omega$} & {IM-SRG(2)} & {IM-SRG(2)$_{R\rightarrow \infty}$} & {DMC} \\
    \midrule
    6  & 1.0  &  20.1583  &  20.1468 &  20.1593(1) \\
       & 0.5  &  11.7690  &  11.7631 &  11.7848(1) \\
       & 0.28 &   7.5713  &   7.5695 &   7.6002(1) \\
       & 0.1  &   3.4962  &   3.4977 &   3.5539(1) \\
    \midrule
    12 & 1.0  &  65.7280  &  65.6990 &  65.7001(1) \\
       & 0.5  &  39.1631  &  39.1473 &  39.1596(1) \\
       & 0.28 &  25.6212  &  25.6140 &  25.6358(1) \\
       & 0.1  &  12.2223  &  12.2234 &  12.2698(1) \\
    \midrule
    20 & 1.0  & 155.9737  & 155.9130 & 155.8822(1) \\
       & 0.5  &  93.9219  &  93.8853 &  93.8752(1) \\
       & 0.28 &  61.94362 &  61.9240 &  61.9268(1) \\
       & 0.1  &  29.95263 &  29.9565 &  29.9779(1) \\
    \midrule
    30 & 1.0  & 308.7673  & 308.6053 & 308.5627(2) \\
       & 0.5  & 187.1671  & 186.8340 & 187.0426(2) \\
       & 0.28 & 124.0410  & 123.7171 & 123.9683(2) \\
       & 0.1  &  60.4319  &  60.1158 &  60.4205(2) \\
    \midrule
    42 & 1.0  & 543.3398  & 543.3304 & 542.9428(8) \\
       & 0.5  & 330.8885  & 330.8425 & 330.6306(2) \\
       & 0.28 & 220.0226  & 219.9534 & 219.8426(2) \\
       & 0.1  & 107.7851  & 106.8074 & 107.6389(2) \\
    \midrule
    56 & 1.0  & 880.4159  &          & 879.3986(6) \\
       & 0.5  & 538.8199  &          & 537.353(2)  \\
       & 0.28 & 360.83689 &          & 358.145(2)  \\
    \bottomrule
  \end{tabular}
\end{table}

@@

The discrepeancy between IM-SRG results and the exact result is due to a combination of two factors: inadequate basis set, and operator truncation.
The former can be reduced by increasing the number of shells, and can be further alleviated by extrapolation techniques.

Both IM-SRG and MP2 improve the accuracy the result dramatically over the more primitive HF.  (It may help to add plots *without* HF.)

[[TODO: add FCI, CCSD results to GS energy plots]]

\subsection{Comparison between methods}

\subsubsection{Ground state energy}

@@

There are a few cases where the IM-SRG over-corrects the result, leading to a ground state energy that is lower than the exact result.  This is not unexpected given that, unlike Hartree-Fock, IM-SRG is non-variational in the presence of operator truncations, which cause small unitarity violations.  Overcorrection tends to occur when the number of particles is few, or the frequency is very low (high correlation).  However, MPPT does not do much better in the high frequency regimes as it often converges slower than IM-SRG.

The operator truncations in the evaluation of the commutator is one of the two sources of discrepancy between IM-SRG and exact results.  The other source is that of basis truncation, which limits the size of the Hilbert space that can be explored by the method.  Unlike operator truncation, which is difficult to overcome as it increases the computational cost exponentially, increasing the basis comes a cost of about $O(n^2 N^4)$ for IM-SRG(2) with the White generator,\footnote{This result holds for any generator whose two-body part contains only pphh- and hhpp-type matrix elements.} [[cite Heiko?]] where $N$ is the total number of states in the single-particle basis and $n$ is the number of particles.  In practice, the scaling cost is usually lower as the conserved symmetries allow the Hamiltonian to be partitioned into blocks.

\subsubsection{Addition and removal energies}

@@

\subsection{Rate of convergence}
